---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# lorax

<!-- badges: start -->
[![R-CMD-check](https://github.com/topepo/lorax/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/topepo/lorax/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/topepo/lorax/graph/badge.svg)](https://app.codecov.io/gh/topepo/lorax)
<!-- badges: end -->

The goal of lorax is to help look at different aspects of tree- and rule-based models. 

lorax supports a few APIs:

 - `as.party()` converts trees to the format used by `partykit::ctree()`, mostly because it has an amazing `plot()` method. 
 - `extract_rules()` helps write out the logical paths to the terminal nodes. 
 - `active_predictors()` enumerates which predictors were actually used in a split. 
 - `var_imp()` is a wrapper for any importance method contained in the package. This is a wrapper that enables a common interface to the scores. 
 
Here is a list of which classes have which methods:

```{r}
#| echo: false
library(lorax)

active_predictors_mthd <-
 tibble::tibble(
  method = "active_predictors",
  class = ls(pattern = "^active_predictors", envir = asNamespace("lorax")) |>
   gsub("active_predictors\\.", "", x = _)
 )

extract_rules_mthd <-
 tibble::tibble(
  method = "extract_rules",
  class = ls(pattern = "^extract_rules", envir = asNamespace("lorax")) |>
   gsub("extract_rules", "", x = _) |>
   gsub("^\\.", "", x = _)
 )

party_mthd <-
 tibble::tibble(
  method = "as.party",
  class = ls(pattern = "^as.party", envir = asNamespace("lorax")) |>
   gsub("as.party\\.", "", x = _)
 )

imp_mthd <-
 tibble::tibble(
  method = "var_imp",
  class = ls(pattern = "^var_imp", envir = asNamespace("lorax")) |>
   gsub("var_imp", "", x = _) |>
   gsub("^\\.", "", x = _)
 )

dplyr::bind_rows(
 imp_mthd,
 active_predictors_mthd,
 party_mthd,
 extract_rules_mthd
) |>
 dplyr::filter(method != class & !grepl("^new_", class) & class != "") |>
 dplyr::mutate(here = cli::symbol$tick) |>
 tidyr::pivot_wider(
  id_cols = class,
  names_from = method,
  values_from = here,
  values_fill = cli::symbol$cross
 ) |>
 dplyr::arrange(tolower(class)) |>
 dplyr::mutate(
  as.party = ifelse(class %in% c("rpart", "cforest", "party"), "n/a", as.party),
  var_imp = ifelse(class %in% c("bart", "C5.0"), "n/a", var_imp)
 ) |> 
 knitr::kable()
```

Note that `as.party.rpart()` is in the partykit package and that cforest is made out of party objects.  

## Installation

You can install the development version of lorax like so:

``` r
require(pak)
pak::pak("topepo/lorax)
```

## Example


```{r}
#| label: startup
#| include: false
library(lorax)
library(tibble)
library(ranger)
library(palmerpenguins)
```

```{r}
#| label: ranger-ex
#| results: hide
set.seed(822)
rngr_fit <- ranger(species ~ ., data = penguins, max.depth = 3, num.trees = 10)
```


```{r}
#| label: ranger-party
rngr_party <- as.party(rngr_fit, data = penguins, tree = 1)
rngr_party

plot(rngr_party)
```

```{r}
#| label: ranger-rules
all_rules <- extract_rules(rngr_party, trees = 10)

# An expression
all_rules$rules[[1]]

# Text
all_rules$rules[[1]] |> rule_text()

# Substitutions
new_names <- 
 tribble(
  ~ original, ~ label,
  "flipper_length_mm", "Flipper Length",
  "body_mass_g", "Body Mass"
 )
 
all_rules$rules[[1]] |> rule_text(key = new_names)

# Bullets:
all_rules$rules[[1]] |> 
 rule_text(key = new_names, bullets = TRUE) |> 
 cat()
```

## Code of Conduct
  
Please note that the lorax project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/1/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
